---
title: "Question 9 Power Problem Set"
author: "Karla Garibay Garcia"
date: "6/9/2022"
output: html_document
---

## Preliminaries

Get R package *DeclareDesign* at declaredesign.org

```{r load, echo=TRUE, warning=FALSE}

rm(list=c()) .
library(DeclareDesign)
library(truncnorm) #for truncated distribution
library(knitr)
library(ggplot2)
library(kableExtra)
```

## declare_population()



```{r population, echo=TRUE}

set.seed(228) #Setting seed for randomization
population <- declare_population(
  waterways = add_level(N=250, 
    water_quality=runif(n=N, min=3.92, max=8.21), 
    u=rnorm(n=N, mean=1.1, sd=0.1)) 
)
```


## Population descriptives

```{r population-see, echo=TRUE, fig.height=5.5}
pop <- population() 
hist(pop[,2], xlab="Baseline Water Quality", 
     main="Baseline", cex=24) 
```

## declare_potential_outcomes()

The next step is to declare the full schedule of potential outcomes $Y(1)$ and $Y(0)$ under an assumption about the effect size of interest.

- Recall that in Jayachandran et al. (2017), tree cover decreased by 13.4 ha in the control group and 7.9 ha in the treatment group.

```{r po, echo=TRUE}
eff_size <- -0.19 
potential_outcomes <- 
  declare_potential_outcomes(
    Y_D_0=water_quality*u,
    Y_D_1=water_quality*u+eff_size) 
```


## Potential outcomes descriptives

```{r po-see, echo=TRUE}
po <- potential_outcomes(pop) 
kable(po[1:5,], digits=1)
```


## declare_sampling()

Next, we want to select the sample size. Let's start with 100 waterways 

```{r sample, echo=TRUE}
samp_size <- 100 
sampling <- 
  declare_sampling(S = draw_rs(N=N, n=samp_size)) 
sam <- sampling(po) 
kable(sam[1:5,c(1:2,4:6)], row.names = FALSE,
      digits = 1)
```


## declare_assignment()


```{r assign, echo=TRUE}
assigning <- 
  declare_assignment(D = conduct_ra(N=N, prob=0.5)) 
assigned <- assigning(sam) 
prop.table(table(assigned$D)) 
```


## Assessing balance


```{r violin, echo=FALSE, fig.height=6}
ggplot(data=assigned, aes(x=as.factor(D), y=water_quality)) +
geom_violin(aes(fill=as.factor(D), color=as.factor(D))) +
theme_minimal(base_size = 24) + xlab("Assignment")
```


## declare_reveal()

This step declares how the potential outcomes are revealed by the random assignment

```{r reveal, echo=TRUE}
revealing <- declare_reveal(assignment_variables=D) 
```

## declare_inquiry()


```{r estimand, echo=TRUE}
estimand <- declare_inquiry(ATE = eff_size)
estimand(po)
```

## declare_estimator()

Next, we declare the estimators we use for recovering the estimand. While there are many advanced estimators, we'll focus on the two core experimental estimators:
1. difference-in-means
2. difference-in-differences

```{r estimator, echo=TRUE}
dim <- declare_estimator(Y ~ D, 
                         inquiry = estimand,
                         model = difference_in_means, 
                         label = "DIM")

did <- declare_estimator(Y - water_quality ~ D, 
                         inquiry = estimand,
                         model = difference_in_means, 
                         label = "DID")
```


## declare_design()

This function brings all of the parts of the process together in a single design and allows for each part of the design to be simulated repeatedly.

```{r design, echo=TRUE}
design <- population + potential_outcomes + sampling +
          assigning + revealing + estimand + dim + did 
```


## diagnose_design()

At this stage, we can calculate various features of the design that we have specified.

```{r diagnosis, cache=TRUE}
diagnosis <- diagnose_design(design, sims=1000) 
diagnosis$diagnosands_df[,c(1,3,5,9,11)] %>%
  kable() 

diagnosis
```


## Looking under the hood, DIM

```{r underhood-dim, height=6, echo=FALSE}
sim.out <- diagnosis$simulations 
hist(sim.out$estimate[sim.out$estimator=="DIM"],
     main="Randomization Distribution",
     xlab="Estimates in Realized Experiments",
     xlim=c(-2,2), cex=24)
abline(v=-0.19, lwd=3, col="red") 
```

## Looking under the hood, DID

```{r underhood-did, height=6, echo=FALSE}
sim.out <- diagnosis$simulations 
hist(sim.out$estimate[sim.out$estimator=="DID"],
     main="Randomization Distribution",
     xlab="Estimates in Realized Experiments",
     xlim=c(-0.7,0.3), cex=24)
abline(v=-0.19, lwd=3, col="red")
```

## modify_design()


```{r more-sample}
samp_size <- 200 #Redefining our sample size to 200 waterways
sampling2 <- 
  declare_sampling(S = draw_rs(N=N, n=samp_size)) 
design2 <- population + potential_outcomes + sampling2 +
          assigning + revealing + estimand + dim + did 
```

## diagnose_design()

Diagnosing the design with twice the sample size

```{r diagnosis2}
diagnosis2 <- diagnose_design(design2, sims=1000) 
diagnosis2$diagnosands_df[,c(1,3,5,9,11)] %>%
  kable()

diagnosis2
```


## Looking under the hood, DIM

```{r underhood-dim2, height=6, echo=FALSE}
sim.out <- diagnosis2$simulations
hist(sim.out$estimate[sim.out$estimator=="DIM"],
     main="Randomization Distribution",
     xlab="Estimates in Realized Experiments",
     xlim=c(-2.5,2), cex=24)
abline(v=-0.19, lwd=3, col="red")
```

## Looking under the hood, DID

```{r underhood-did2, height=6, echo=FALSE}
sim.out <- diagnosis2$simulations
hist(sim.out$estimate[sim.out$estimator=="DID"],
     main="Randomization Distribution",
     xlab="Estimates in Realized Experiments",
     xlim=c(-0.7,0.3), cex=24)
abline(v=-0.19, lwd=3, col="red")
```

## redesign()
What is the minimum sample size required to conduct a sufficiently powered evaluation using each estimator?
```{r other-sample-sizes1}
diagnoses <- design %>% 
  redesign(samp_size=seq(20,250,20)) %>% 
  diagnose_design(sims=150)

diagnoses # Neither DID or DIM achieve a power of 0.8 or 0.95. The highest power achieved by DID is 0.67 at a sample size of 240 and by DIM is 0.21 at a sample size of 200. 
```

## redesign()
What is the minimum sample size required to conduct a sufficiently powered evaluation using each estimator?
```{r other-sample-sizes2, height=6, echo=FALSE}
diagnoses$diagnosands_df %>% 
  ggplot(aes(x=samp_size,y=power)) +
  geom_hline(aes(yintercept=1), lty="dotted", col="red", size=0.5) +
  geom_point() +
  scale_y_continuous(limits=c(0, 2)) +
  theme(text=element_text(size=13)) +
  facet_grid(.~estimator)
```


## redesign()
How strong does treatment need to be for our evaluation to be sufficiently powered, given a fixed sample size of 100?
```{r other-mde1}
diagnoses2 <- design %>% 
  redesign(eff_size=seq(-0.9, 0.9, 0.05)) %>% 
  diagnose_design(sims=250)

diagnoses2
```

## redesign()
How strong does treatment need to be for our evaluation to be sufficiently powered, given a fixed sample size of 100?
```{r other-mde2, height=6, echo=FALSE}
diagnoses2$diagnosands_df %>% # trouble here
  ggplot(aes(x=eff_size,y=power)) +
  geom_hline(aes(yintercept=0.8), lty="dotted", col="red", size=0.5) +
  geom_point() +
  scale_y_continuous(limits=c(0, 2)) +
  theme(text=element_text(size=13)) +
  facet_grid(.~estimator)
```


